\vspace*{-8mm}\textbf{\textsc{Novelty Detection for Semantic Place Categorization}}

### 2011-06-24 thesis/acknows.tex
Semantic mapping missing points and other small fixes.

Virgile Högman, for besides handling me as a work colleague and chatter-box, becoming a friend.
And Erik Ass for all the interesting coding and math discussions on totally random problems.
I can't also avoid to thank my neighbours back in Nockeby, in special to Diogo Gonçalves and Elise Löbker
that kept chatting with me through the nights during my thesis work.

Thanks to Mariatorget people who kept asking how my thesis was going and when it was going to be finished.
And for all the parties, beer relaxing time and nights out! Thanks to Elina Säfsten, for being a cool neighbour
with whom I had the pleasure to spent quite some hours talking and partying with her and her Swedish friends.
Thanks to Manuel Gattermayr and Tommaso Facchini for being awesome neighbours as well.

I am also grateful to Bernhard Schwaighofer, Manuel and Tommaso and others for an amazing
road trip, night fires, sunsets and sunrises on Sweden. Special thanks to Bernie for during the road trip
promising to wake me up before my sleeping bag starts to burn and melt with my skin and eventually turning me water proof.

@@

\vspace{10mm}
\flushleft{
"I am thankful to all the awesome people who were part of this Stockholm chapter of my life"
--
André Susano Pinto}

>>>>

Virgile Högman, for besides handling me as a work colleague and chatter-box, becoming a friend.
And Erik Ass for all the interesting coding and math discussions on totally random problems.
I cannot also avoid to thank my neighbours back in Nockeby, in special to Diogo Gonçalves and Elise Löbker
that kept chatting with me through the nights during my thesis work.

Thanks to Mariatorget people who kept asking how my thesis was going and when it was going to be finished.
And for all the parties, drinking, relaxing and nights out! Thanks to Elina Säfsten, for being a cool neighbour
with whom I had the pleasure to spent quite some time talking and partying with her and her Swedish friends.
Thanks to Manuel Gattermayr and Tommaso Facchini for being awesome neighbours as well.

I am also grateful to Bernhard Schwaighofer, Manuel, Tommaso and others for an amazing
road trip, night fires, sunsets and sunrises on Sweden. Special thanks to Bernie for during the road trip
promising to wake me up before my sleeping bag starts to burn and melt with my skin and eventually turning me water proof.

@@

\vspace{10mm}
\flushleft{
"I am thankful to all the awesome people who were part of this Stockholm chapter of my life"
--
\emph{André Susano Pinto}}

### 2011-06-24 thesis/semantic-mapping.tex
moved line in place. >.<
Semantic mapping missing points and other small fixes.

% motivation
Humans associate concepts to areas that relate to the functionality of those. That
association is either useful by explaining where to find certain objects as well to
understand the types of activities to perform on it. 
Those human-concepts are meant to be meaningful for humans and by incorporating conceptual
knowledge about them it is expected to improve the agent ability to interact and communicate
with humans and the ability perform human-like tasks.

% semantic mapping definition
In this context semantic mapping is seen as the process of building a representation of
the environment which associates spatial entities with a set of defined spatial concepts (in this case the human concepts).

Missing:
\begin{itemize}
\item importance of conceptual knowledge
% In order to build such a mapping of the environment, it is important to define the knowledge that allows the agent to reason
% about.

\item application of conceptual knowledge
% It is expected that by being able to reason over those semantics an agent can better
% understand human environments and increase its efficiency when performing human-tasks on them.
% By handling this conceptual knowledge the agent becomes aware of

\item challenges and need for representing uncertainty
% The goal of a spatial representation is to provide the agent with a view about the environment
% that allows it to reason beyond its sensory horizon.

% Such a representation is by its nature imperfect, incomplete, inaccurate and invalid.

% world that allows the agent to perform reasoning over it and efficiently perform tasks.
% It should also give the agent ability to reason about parts 
% Such a representation cannot be perfect since the methods and information are uncertain

% It is therefore clear that the need to handle information on an high-level semantic and
% probabilistic fashion arises not only for efficiency reasons but also from the desire
% of being able to communicate and reason on those aspects.
% In order to fully explore the semantic properties and their relations it is necessary to
% detect and classify semantic aspects as well to understand the relationships between them.
% That conceptual knowledge is important and any limitations on it will limit the capacities
% of an agent to reason. 
\end{itemize}

% here a here a method is used that does semantic mappping and representes conceptual knowledge
% on a probabilistic fashion.
This thesis motivation lies on developing methods for detecting knowledge gaps on the spatial
semantic knowledge of a mobile agent. For that spatial knowledge needs to be defined, as well
the methods used to perform the semantic mapping.
In that context this chapter presents \gls{Dora}, a system that represents its spatial knowledge
with the representation proposed by \cite{pronobis2010ias}, and performs semantic mapping using
inference on a probabilistic graphical model.

@@

\subsection{Features}
\label{sec:dora-features}
The \emph{categorical layer} requires classifiers with the ability to categorize the
low\hyp{}level features into categories.
For that \gls{Dora} performs detection on the following categorical entities:
\begin{description}
\item[Doorway Detection] by using the laser range finder the system tries to detect doors.
\item[Geometric Shape] by using features extracted from the laser data and pre-trained classifiers
the system tries to classify rooms according to their shapes (e.g.\ rectangular, square).
\item[Geometric Size] is classified once again recurring to laser data. In this case the system
\item[Object Detection] is performed by running traditional object recognition classifiers.
Those are previously trained on a supervised fashion to later be able to detect objects
based on their visual features, e.g.\ by using \gls{SIFT} as described in \autoref{sec:sift}.
tries to categorize the room size according to a set of previously learned concepts (e.g. large, medium, small).
\item[Visual Appearance] is extracted from the visual input. For that Dora uses a trained
a set of classifiers on global visual features such as \gls{CRFH} to categorize the visual
appearance of a room in categories such as corridor, office, library.
\end{description}

>>>>

% motivation
Humans associate concepts to areas that relate to the functionality of those. That
association is either useful by explaining where to find certain objects as well to
understand the types of activities to perform on certain areas. 
Those human-concepts are meant to be meaningful for humans and by incorporating conceptual
knowledge about them it is expected to improve the agent ability to interact and communicate
with humans and to perform human-like tasks.

% semantic mapping definition
In this context semantic mapping is seen as the process of building a representation of
the environment which associates spatial entities with a set of defined spatial concepts (in this case human concepts).

% Importance of spatial and conceptual knowledge
In order to build such a semantic map, it is important to define the spatial and conceptual knowledge
that allows an agent to map between its machine-friendly representation of the world and the high-level semantics
characteristic of the human world.
That knowledge needs to define concepts and how to segment and identify them from the low-level representation,
it needs also to define how the concepts relate to each other and attribute meanings to them.
In the case of indoor environments this can be seen for example as: rooms are a unit of division of area and are
often separated by doors, each room can also be characterized by the objects inside, size and shape, and they
can be categorized in meaningful concepts such as: kitchen, library, corridor, etc\dots

% Why to treat it in a probabilistic way
The low-level representation and the knowledge of the agent, will always be to a certain a degree imperfect,
incomplete, inaccurate and invalid. Due to that both the concepts and their relations
cannot be quantified in a rule base system and ought to be treated in a probabilistic way in order to allow
the agent to still reason with some accuracy about the world.

% here a here a method is used that does semantic mapping and represents conceptual knowledge
% on a probabilistic fashion.
This thesis motivation lies on developing methods for detecting knowledge gaps on the spatial
semantic knowledge of a mobile agent. For that spatial knowledge needs to be defined, as well
the methods used to perform the semantic mapping.
In that context this chapter presents \gls{Dora}, a system that represents its spatial knowledge
with the representation proposed by \cite{pronobis2010ias}, and performs semantic mapping using
inference on a probabilistic graphical model.

@@

\subsection{Features}
\label{sec:dora-features}
The \emph{categorical layer} requires classifiers with the ability to categorize the
low\hyp{}level features into categories.
For that \gls{Dora} performs detection on the following categorical entities:
\begin{description}
\item[Doorway Detection] by using the laser range finder the system tries to detect doors.
\item[Geometric Shape] by using features extracted from the laser data and pre-trained classifiers
the system tries to classify rooms according to their shapes (e.g.\ rectangular, square).
\item[Geometric Size] is classified once again recurring to laser data. In this case the system
tries to categorize the room size according to a set of previously learned concepts (e.g. large, medium, small).
\item[Object Detection] is performed by running traditional object recognition classifiers.
Those are previously trained on a supervised fashion to later be able to detect objects
based on their visual features, e.g.\ by using \gls{SIFT} as described in \autoref{sec:sift}.
\item[Visual Appearance] is extracted from the visual input. For that Dora uses a trained
a set of classifiers on global visual features such as \gls{CRFH} to categorize the visual
appearance of a room in categories such as corridor, office, library.
\end{description}
